---
# title: "`R` in Action"
# subtitle: "Efficient data science with `R`"
format:
    revealjs:
        width: 1280
        height: 720
        pagetitle: "Research fair 2024"
        menu: false
        history: false
        transition: slide
        background-transition: slide
        controls: true
        controls-layout: bottom-right
        navigation-mode: linear
        # theme: serif
        # center: true
        fig-align: center
        auto-slide: 5000
        auto-slide-stoppable: true
        loop: true
        embed-resources: true
editor_options: 
  chunk_output_type: console
---

## `R` in Action {.center}
::: {.r-fit-text}
Efficient data science with `R`
:::
A demonstration by Md. Aminul Islam Shazid


## {.center data-autoslide="5000"}
::: {.r-fit-text}
Grammar of graphics with `ggplot2`
:::


## Plots using grammar of graphics with `ggplot2` {data-autoslide="15000"}
- `ggplot2` is an `R` package that implements the grammar of graphics.
- Can provide beautiful graphics with some simple building blocks.
- Variables/features/columns are mapped to various elements of the plot called "aesthetics", e.g., axis, colours, point size, line type etc.
- Then a geometry transforms that "aesthetic" mapping into a plot.

## A simple example
```{r}
library(ggplot2)
library(palmerpenguins)
data(package = 'palmerpenguins')
# attributes(penguins$bill_length_mm)$label <- "Bill length (mm)"
```

```{r}
#| echo: true
#| warning: false
ggplot(data = penguins, 
       mapping = aes(x = bill_length_mm, y = flipper_length_mm)) + 
    geom_point()
```

## Adding a grouping variable
```{r}
#| echo: true
#| warning: false
ggplot(penguins, 
       mapping = aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
    geom_point()
```

## Let's add another dimension to the plot!
```{r}
#| echo: true
#| warning: false
ggplot(penguins, 
       mapping = aes(x = bill_length_mm, y = flipper_length_mm, color = species, 
                     size = body_mass_g)) + 
    geom_point(alpha = 0.5)
```

## Adding yet another dimension!
```{r}
#| echo: true
#| warning: false
ggplot(penguins, 
       mapping = aes(x = bill_length_mm, y = flipper_length_mm, color = species, 
                     size = body_mass_g)) + 
    geom_point(alpha = 0.5) +
    facet_wrap(~island)
```

## Comparing a variable across groups with boxplot
```{r}
#| echo: true
#| warning: false
ggplot(penguins,
       mapping = aes(y = body_mass_g, x = species, fill = species)) +
    geom_boxplot(show.legend = F)
```

```{r}
# geom_boxplot(position = position_dodge2(preserve = "single"))
```


## Violon plots as alternative to boxplot
More informative: gives a sense of the density too!
```{r}
#| echo: true
#| warning: false
ggplot(penguins,
       mapping = aes(y = body_mass_g, x = species, fill = species)) +
    geom_violin(show.legend = F)
```



## {.center data-autoslide="5000"}
::: {.r-fit-text}
Publication ready tables with `gtsummary`
:::


## A title

```{r}
#| output-location: column
#| results: hold
#| echo: true
plot(mtcars$mpg, mtcars$wt)
```

## Another title {auto-animate="true"}

test

## Another title {auto-animate="true"}

Another test
